% Chapter 14 — Implementation Quality and Testing

\chapter{Implementation Quality \& Testing}
\label{ch:qa}

\section{Automated tests}
The backend ships with Jest + Supertest-style tests covering authentication, courses, enrollments, quizzes, and analytics endpoints where implemented. Automated tests catch regressions when schema or controllers change.

The frontend uses Vitest + Testing Library for representative UI flows (example: course list smoke tests). Tests do not replace manual QA but shorten feedback loops during assignment deadlines.

\section{Manual verification matrix}
Use Table~\ref{tab:manualqa} during demos. Replace Pass/Fail with your actual results.

\begin{longtable}{@{}p{4.2cm}p{7.8cm}p{2.2cm}@{}}
\caption{Manual verification checklist}\label{tab:manualqa}\\
\toprule
\textbf{Scenario} & \textbf{What ``good'' looks like} & \textbf{Result} \\
\midrule
Fresh \texttt{ddl.sql} run & All tables, indexes, views, triggers create without errors & Pass / Fail \\
Create course + content & Rows appear in \texttt{courses} and \texttt{content}; FK chain valid & Pass / Fail \\
Duplicate enrollment & Second enroll rejected or safely ignored per API contract & Pass / Fail \\
Quiz submit & Score matches manual calculation from correct options & Pass / Fail \\
Refresh analytics & Materialized views update after RPC / admin action & Pass / Fail \\
Login / logout & Tokens invalidated appropriately; protected routes reject missing JWT & Pass / Fail \\
\bottomrule
\end{longtable}

\section{Operational hygiene}
\begin{itemize}[leftmargin=*]
  \item Keep development and production credentials separate (\texttt{.env} never committed).

  \item Log enough on the API to debug failing inserts without printing secrets.

  \item Validate all mutating payloads (\texttt{express-validator}) so malformed JSON cannot crash controllers.

\end{itemize}

\section{Known coursework angle}
Expect examiners to ask: ``What would you test next?'' Reasonable answers: load testing on \texttt{activity\_log}, backup/restore drill, stricter RLS policies if mobile clients hit Supabase directly.
