% Chapter 15 — Deployment Considerations

\chapter{Deployment Considerations}
\label{ch:deploy}

\section{Typical hosting split}
LearnTrack is commonly deployed with:

\begin{itemize}[leftmargin=*]
  \item \textbf{Vercel} for the React build (static assets + SPA routing).

  \item \textbf{Render} (or similar) for the Node Express API.

  \item \textbf{Supabase} for PostgreSQL, Auth, and optional storage.

\end{itemize}

\section{Environment variables that must match}
\begin{itemize}[leftmargin=*]
  \item \texttt{VITE\_API\_URL} on the frontend must point to the deployed API \textbf{including} \texttt{/api/v1} if your client code expects that base.

  \item \texttt{CLIENT\_ORIGIN} (or equivalent CORS setting) on the backend must exactly match the deployed frontend URL---including \texttt{https} and no trailing slash mismatches.

  \item Database URL and service role key live only on the server; never embed them in React.

\end{itemize}

\section{Common production mistakes (and fixes)}
\begin{itemize}[leftmargin=*]
  \item \textbf{404 on all routes:} API base URL missing \texttt{/api/v1} segment.

  \item \textbf{CORS errors in browser console:} backend allowed origin does not match Vercel domain.

  \item \textbf{Stale frontend env vars:} Vite bakes \texttt{VITE\_*} at build time---redeploy after changing them.

  \item \textbf{Cold starts on free tiers:} first request after idle may be slow; retry logic or warm-up ping helps demos.

\end{itemize}

\section{Demo checklist}
\begin{enumerate}[leftmargin=*]
  \item Hit \texttt{/api/v1/health} and confirm JSON status.

  \item Log in from deployed UI.

  \item Create a course and one quiz; submit one attempt as a student.

  \item Trigger analytics refresh if instructor tiles look empty.

  \item Confirm browser console has no CORS errors.

\end{enumerate}
