% Chapter 16 — Conclusion

\chapter{Conclusion}
\label{ch:conclusion}

\section{What we demonstrated}
LearnTrack is a complete learning platform backed by a relational database. Students generate real telemetry (activity and progress). Instructors author structured courses and quizzes. Administrators can oversee users and global metrics depending on implemented routes.

From a DBMS course perspective, the project shows:

\begin{itemize}[leftmargin=*]
  \item \textbf{Modeling:} entities, relationships, lookup tables, and clear keys.

  \item \textbf{Integrity:} foreign keys, uniqueness, check constraints, soft deletes where history matters.

  \item \textbf{Analytics:} materialized views that precompute expensive joins for dashboards.

  \item \textbf{Programmability:} triggers and functions for timestamps, instructor rules, and bulk refresh.

  \item \textbf{Integration:} a modern stack (React + Express + Supabase) where SQL remains the source of truth.

\end{itemize}

\section{Limitations (honest)}
\begin{itemize}[leftmargin=*]
  \item Materialized views can lag minutes behind raw logs unless refreshed often.

  \item Very large \texttt{activity\_log} tables eventually need partitioning or archival---outside typical coursework scope.

  \item Row Level Security exists in DDL but the report emphasises API-level auth for clarity.

\end{itemize}

\section{Future extensions}
\begin{itemize}[leftmargin=*]
  \item Scheduled refresh jobs with logged timestamps per view.

  \item Certificate generation stored as immutable rows.

  \item Read replicas or ETL export if analytics traffic grows.

\end{itemize}

\section{Closing statement}
The database is not a filing cabinet---it is the contract that keeps learning data trustworthy. LearnTrack shows how normalized tables, constraints, and targeted summaries work together so teaching teams can see engagement without sacrificing correctness.

\begin{table}[htbp]
\centering
\caption{Summary: what the DBMS project delivered}
\begin{tabular}{@{}p{3.7cm}p{8.9cm}@{}}
\toprule
\textbf{Area} & \textbf{Outcome} \\
\midrule
Model & Normalized schema with lookup tables and explicit relationships. \\
Integrity & Constraints prevent duplicates and orphan rows in normal use. \\
Analytics & Materialized views speed up instructor and admin dashboards. \\
Automation & Triggers/functions keep profiles and timestamps consistent. \\
Deployment & End-to-end operation possible on hosted frontend + API + Supabase. \\
\bottomrule
\end{tabular}
\end{table}
