% Appendices

\appendix

\chapter{REST Resource Inventory (Prefix \texttt{/api/v1})}
\begin{itemize}[leftmargin=*]
  \item \textbf{Health:} \texttt{GET /health} --- simple alive check.

  \item \textbf{Auth:} \texttt{/auth/register}, \texttt{/auth/login}, \texttt{/auth/logout} --- identity and session lifecycle.

  \item \textbf{Users:} profile and admin user management routes as implemented in \texttt{users.js}.

  \item \textbf{Courses:} public catalogue, instructor ``mine,'' single course with nested content and quizzes, CRUD, roster.

  \item \textbf{Enrollments:} list mine, enroll, status updates, admin deletes where enabled.

  \item \textbf{Activity \& progress:} log events, read/update completion percentages.

  \item \textbf{Quizzes:} create/update questions, submit attempts, review attempts.

  \item \textbf{Analytics:} active students, underperforming learners, skipped content, completion rates, instructor course dashboard, admin dashboard, optional materialized view refresh.

\end{itemize}

\chapter{Mini Glossary}
\begin{description}[style=nextline]
  \item[Foreign key]
  A column that must match an existing primary key in another table---keeps relationships honest.

  \item[Materialized view]
  Saved result of a query refreshed on demand or schedule; faster reads, slightly stale data.

  \item[Soft delete]
  Mark a row inactive (\texttt{deleted\_at}) instead of removing it, preserving references.

  \item[Surrogate key]
  Database-generated ID (e.g.\ SERIAL) used purely to identify a row.

  \item[UUID]
  Global unique identifier for users here aligned with Supabase Auth.

\end{description}

\chapter{Repository File Map}
\begin{itemize}[leftmargin=*]
  \item \texttt{backend/src/db/ddl.sql} --- full schema (tables, indexes, views, triggers, RLS).

  \item \texttt{backend/src/db/seed.sql} --- optional demo data.

  \item \texttt{backend/src/db/connection.js} --- Supabase client bootstrap.

  \item \texttt{backend/src/controllers/*.js} --- API logic and queries.

  \item \texttt{learntrack-client/src/api/index.js} --- Axios base URL and endpoint helpers.

  \item \texttt{latex/chapters/*.tex} --- modular chapters for this report (included from \texttt{DBMS\_PROJECT\_REPORT.tex}).

\end{itemize}
